|

### 2019-05-29 manuscript/article/article.Rmd
update manuscript

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "We thank Umberto Mignozzetti, Lucas Mingardi and Robert Myles McDonnell for their helpful comments. Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
abstract: "This article presents a georeferenced dataset on human rights violations in the Pinochet dictatorship in Chile. We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations. The dataset comprises XXXXXXXXXX variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`. As examples, we describe the major temporal and spatial patterns of the human rights abuses. We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential applications of the dataset."
keywords: "Chile, georeferenced event data, human rights, Pinochet, truth commission"
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    template: template.latex
---

@@

Our dataset comprises 2,398 observations and XXXXXXXXX variables. Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number. There are several variables describing the personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available. The data also feature geographical coordinates for a number of the incidents, as well as information about methods of torture and causes of death. Table 1 shows an example entry from our dataset.

@@

| Variable                | Sample          |
| :---------------------- | :-------------- |
| id                      | 276             |
| page                    | 265             |
| start date              | 18/01/1974      |
| end date                | 18/01/1974      |
| last name               | Marques Agusto  |
| first name              | Nelson José     |
| minor                   | 0 (no)          |
| age                     | 31              |
| male                    | 1 (yes)         |
| occupation              | office worker   |
| victim affiliation      | communist party |
| violence                | killed          |
| method                  | gun             |
| interrogation           | NA (unknown)    |
| torture                 | NA (unknown)    |
| mistreatment            | 1 (yes)         |
| perpetrator             | military        |
| place                   | in public       |
| start location          | Iquique         |
| latitude 1              | -20.23070       |
| longitude 1             | -70.13567       |
| place 2                 | in custody      |
| location 2              | Pisagua prison  |
| latitude 2              | -19.59667       |
| longitude 2             | -70.21222       |

Table: Sample event from the dataset (selected variables).

@@

| Violence         | _N_    |
| :--------------- | -----: |
| Killed           |   1331 |
| Disappeared      |    853 |
| Unresolved cases |     93 |
| Suicides         |      7 |
| Unknown          |    115 |

Table: Number of killings, disappearances, unresolved cases, and suicides.

@@

\vspace{.5cm}
\begin{center}
Figure 1 about here.

Two time trends: one with the whole period,\\ another discarding the first year to better visualise variation in later years
\end{center}
\vspace{.5cm}

[comment]: <> (Figure 01 - Time Trends Here)
\newpage

In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals as "definite and known deaths". The violations continued in October, with 504 incidences and 317 certified deaths. The numbers decline to about 10% of those levels in November and December, and then recede further in 1974.

@@

@@

\vspace{.5cm}
\begin{center}
Figure 2 about here.

Two maps: deaths and disappearances 
\end{center}
\vspace{.5cm}

[comment]: <> (Figure 02 - Spatial variation)

We can further disaggregate the analysis and visualise how the regime persecuted particular groups. Students and members of left-wing parties are two categories of special interest. On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium. Military forces tortured and murdered hundreds of prisoners over the course of a few days. An emblematic victim was Víctor Lidio Jara Martínez, a university teacher, theatre director and singer (id number 25). According to witnesses, Mr Jara had his hand crushed and his wrists broken, and he was later shot 44 times with a machine gun. Mr Jara's body was left in the front of the stadium as a warning for other political prisoners [@mcsherry2015victor 56]. The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR). The MIR originated in student unions and in the late 1970s the movement promoted a series of attacks against government personnel and official buildings. Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].

\vspace{.5cm}
\begin{center}
Figure 3 about here.

Time trends: attacks against students and workers
\end{center}
\vspace{.5cm}

>>>>

---
title:  |
        | **Deaths and Disappearances in the Pinochet Regime: A New Dataset**
author: 
  - "Danilo Freire[^df]"
  - "John Meadowcroft[^jm]"
  - "David Skarbek[^ds]"
  - "Eugenia Guerrero[^eg]"
date: \today
thanks: "We thank Umberto Mignozzetti, Lucas Mingardi and Robert Myles McDonnell for their helpful comments. Replication materials and R source code are available at [`http://github.com/danilofreire/pinochet`](http://github.com/danilofreire/pinochet)."
abstract: "This article presents a georeferenced dataset on human rights violations in the Pinochet dictatorship in Chile. We coded the personal details of 2,398 victims named in the Chilean Truth Commission Report and added geographical coordinates for all identifiable atrocity locations. The dataset comprises 59 variables from 1973 to 1990 and is available as a stand-alone spreadsheet or as the `pinochet` package for `R`. As examples, we describe the major temporal and spatial patterns of the human rights abuses. We also discuss our coding procedures, show descriptive statistics, and conclude with suggestions for potential applications of the dataset."
keywords: "Chile, georeferenced event data, human rights, Pinochet, truth commission"
fontfamily: libertine
fontawesome: yes
fontsize: 12pt
monospace-url: yes
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: yes
    template: template.latex
---

@@

Our dataset comprises 2,398 observations and 59 variables. Each observation corresponds to a victim of the Pinochet regime and every individual has a unique id number. There are several variables describing the personal information of the victims, such as age, gender, nationality, occupation, and political affiliation if available. The data also feature geographical coordinates for a number of the incidents, as well as information about methods of torture and causes of death. Table 1 shows an example entry from our dataset.

@@

```{r sample-obs, echo=FALSE, warning=FALSE, message=FALSE,cache=TRUE}
if (!require("kableExtra")) {
        install.packages("kableExtra")
}
if (!require("tidyverse")) {
        install.packages("tidyverse")
}
load("/Users/politicaltheory/Documents/github/pinochet/data/pinochet.RData")
kable(t(pinochet[12, c(1:4,7:22,24:37)]), "latex", caption = "Sample event from the dataset (selected variables)", col.names = c("Value"), booktabs = T) %>%
  kable_styling(position ="center", latex_options = "hold_position")
```

\vspace{.5cm}

@@

\vspace{.5cm}

```{r number, echo=FALSE, message=FALSE, warning=FALSE,cache=TRUE}
kable(table(pinochet$violence), "latex", caption = "Number of killings, disappearances, unresolved cases, and suicides", booktabs = T, col.names = c("Violence", "N")) %>%
  kable_styling(position ="center", latex_options = "hold_position")
```

\vspace{.5cm}

@@

In September 1973 alone, the government committed 633 acts of violence against civilians, of which 406 the Commission signals as "definite and known deaths". The violations continued in October, with 504 incidences and 317 certified deaths. In total, the Pinochet regime made 1274 victims in its first year. The numbers decline to about 10% of those levels in November and December, and then recede further in 1974.

@@

\vspace{.5cm}

```{r time-trend, echo=FALSE, fig.align='center', fig.height=3, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Human rights abuses in the Pinochet regime, 1973-1990',cache=TRUE}
if (!require("tidyverse")) {
        install.packages("tidyverse")
}
if (!require("lubridate")) {
        install.packages("lubridate")
}
pinochet %>% ungroup() %>% 
  mutate(Year = year(start_date_monthly)) %>%
  group_by(Year) %>% tally() %>% 
  filter(!is.na(Year)) %>% 
  ggplot(aes(x = Year, y = n)) +
  geom_line() +
  theme_minimal() +
  labs(x = NULL, y = NULL, title = "Number of Human Rights Abuses", 
       subtitle = "Pinochet Regime, 1973-1990") +
  scale_y_continuous(breaks = c(0, 500, 1000, 1274))
```

\vspace{.5cm}

@@

```{r maps, echo=FALSE, warning=FALSE, message=FALSE, fig.align='center', fig.cap="Spatial variation in human rights abuses in the Pinochet regime, 1973-1990", fig.height=6, fig.width=7,cache=TRUE}
if (!require("sf")) {
        install.packages("sf")
}
if (!require("devtools")) {
        install.packages("devtools")
}
devtools::install_github("ropensci/rnaturalearthdata")
if (!require("rnaturalearthhires")) {
        install.packages("rnaturalearthhires", repos = "http://packages.ropensci.org", type = "source")
}
library(rnaturalearthdata)
library(rnaturalearthhires)

chile <- rnaturalearthhires::countries10 %>%
st_as_sf() %>%
filter(SOVEREIGNT %in% c("Chile", "Argentina", "Peru", "Paraguay"))

violent_events <- pinochet %>% 
  select(violence, latitude_1, longitude_1) %>%
  filter(!is.na(latitude_1), !is.na(longitude_1), !is.na(violence)) %>%
  st_as_sf(coords = c("longitude_1", "latitude_1"), crs = 5361)

coords_vio <- st_coordinates(violent_events) %>% as_tibble()
violent_events <- bind_cols(violent_events, coords_vio)

ggplot() +
  geom_sf(data = chile) +
  coord_sf(xlim = c(-75.6, -67), ylim = c(-55, -19)) +
  labs(x = NULL, y = NULL) +
  geom_point(data = violent_events, aes(x = X, y = Y, colour = violence, fill = violence), shape = 21) + 
  scale_colour_manual(values = c("#042333FF", "#481567FF",  "#253582FF", "#2D708EFF", "#29AF7FFF")) +
  scale_fill_manual(values = alpha(c("#042333FF", "#481567FF", "#253582FF", "#2D708EFF", "#29AF7FFF"), .6)) +
  facet_wrap(~violence, nrow = 1) +
  theme_minimal() +
  theme(strip.text = element_blank()) +
  theme(legend.position = "bottom",
        axis.text = element_blank()) + 
  theme(legend.title=element_blank())
```

We can further disaggregate the analysis and visualise how the regime persecuted particular groups. School and university students are two categories of special interest. On 18 September 1073, only one week after the coup, Pinochet and his followers rounded up thousands of students and activists in Chile's National Stadium. Military forces tortured and murdered dozens of prisoners over the course of a few days. The government again tortured and killed students and activists in its campaign against the Revolutionary Left Movement (_Movimiento de Izquierda Revolucionaria_, MIR). The MIR originated in student unions and in the late 1970s the movement promoted a series of attacks against government personnel and official buildings. Government forces strongly retaliated and the group resorted to terrorist tactics to confront the regime [@schlotterbeck2018beyond].

\vspace{.5cm}

```{r time-trend-students, echo=FALSE, fig.align='center', fig.height=3, fig.width=5, warning=FALSE, message=FALSE, fig.cap='Human rights abuses against students, 1973-1990',cache=TRUE}

pinochet %>% ungroup() %>% 
  mutate(Year = year(start_date_monthly)) %>%
  group_by(Year) %>% tally() %>% 
  filter(!is.na(Year)) %>% 
  ggplot() +
  geom_line(aes(x = Year, y = n)) +
  geom_line(aes(x = Year, y = n)) +
  theme_minimal() +
  labs(x = NULL, y = NULL, title = "Number of Human Rights Abuses Against Students", 
       subtitle = "Pinochet Regime, 1973-1990") +
  scale_y_continuous(breaks = c(0, 500, 1000))
```

\vspace{.5cm}
